Solar wind prediction for the Parker Solar Probe orbit

### 2017-07-16 paperMVVB/sections/abstract.tex
replaced 'exponential' with 'power law'

\abstract
{\color{Green4} In view of the planned near-Sun spacecraft mission Parker~Solar~Probe (PSP) (formerly Solar~Probe~Plus) the solar wind environment for its prime mission duration (2018--2025) and down to its intended closest perihelion (\num{9.86}~solar radii) is extrapolated using in situ data. The PSP mission will be humanity's first in situ exploration of the solar corona. Visiting this yet uncharted region is of special interest, because it will help answer hitherto unresolved questions on the heating of the solar corona and the source and acceleration of the solar wind and solar energetic particles. The solar wind extrapolation of this study is performed within the project CGAUSS (Coronagraphic German And US Solar~Probe Survey) which is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR).}	%context
{\color{Green4} We present an empirical solar wind model for the inner heliosphere which is derived from Helios and OMNI in situ data. The German-US space probes Helios~1 and Helios~2 flew in the 1970s and observed solar wind in the ecliptic within heliocentric distances of \SIrange{0.29}{0.98}{\au}. The OMNI database at NASA's Space Physics Data Facility (SPDF) consists of multi-spacecraft intercalibrated in situ data obtained near \SI{1}{\au}. The solar wind model is used together with sunspot number predictions to estimate the frequency distributions of major solar wind parameters PSP will encounter during its mission.}	%aims
{\color{Green4} The model covers the solar wind's magnetic field strength and its plasma parameters proton density, velocity and temperature.
Their individual frequency distributions are represented with lognormal functions. In addition, we also consider the velocity distribution's bi-componental shape, consisting of a slower and a faster part. The model accounts for solar activity and for solar distance dependency \hl{by shifting of these lognormal distributions.} We compile functional relations to solar activity by correlating and fitting the frequency distributions with the sunspot number (SSN), using almost five solar cycles of OMNI data. Further, based on the combined data set from both Helios probes, the parameters' frequency distributions are fitted with respect to solar distance to obtain exponential dependencies. Finally, by combining the found solar cycle and solar distance relations, we obtain a simple dynamical solar wind model for the inner heliosphere, confined to the ecliptic region.}	%methods
{\textcolor{Green4}{ The inclusion of SSN predictions and the extrapolation to the PSP perihelion region enables us to estimate the solar wind environment for PSP's planned trajectory during its prime mission duration 2018--2025.} \hl{The estimated solar wind values at PSP's nearest perihelion are: ... Their value vary up to} \SI{+-0.0000}{\percent}, \hl{arising only from differing amplitude assumptions for the next solar cycle.}}	%results
{}	%conclusions

>>>>

\abstract
{\color{Green4} In view of the planned near-Sun spacecraft mission Parker~Solar~Probe (PSP) (formerly Solar~Probe~Plus) the solar wind environment for its prime mission duration (2018--2025) and down to its intended closest perihelion (\num{9.86}~solar radii) is extrapolated using in situ data. The PSP mission will be humanity's first in situ exploration of the solar corona. Visiting this yet uncharted region is of special interest, because it will help answer hitherto unresolved questions on the heating of the solar corona and the source and acceleration of the solar wind and solar energetic particles. The solar wind extrapolation of this study is performed within the project CGAUSS (Coronagraphic German And US Solar~Probe Survey) which is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR).}	%context
{\color{Green4} We present an empirical solar wind model for the inner heliosphere which is derived from Helios and OMNI in situ data. The German-US space probes Helios~1 and Helios~2 flew in the 1970s and observed solar wind in the ecliptic within heliocentric distances of \SIrange{0.29}{0.98}{\au}. The OMNI database at NASA's Space Physics Data Facility (SPDF) consists of multi-spacecraft intercalibrated in situ data obtained near \SI{1}{\au}. The solar wind model is used together with sunspot number predictions to estimate the frequency distributions of major solar wind parameters PSP will encounter during its mission.}	%aims
{\color{Green4} The model covers the solar wind's magnetic field strength and its plasma parameters proton density, velocity and temperature.
Their individual frequency distributions are represented with lognormal functions. In addition, we also consider the velocity distribution's bi-componental shape, consisting of a slower and a faster part. The model accounts for solar activity and for solar distance dependency \hl{by shifting of these lognormal distributions.} We compile functional relations to solar activity by correlating and fitting the frequency distributions with the sunspot number (SSN), using almost five solar cycles of OMNI data. Further, based on the combined data set from both Helios probes, the parameters' frequency distributions are fitted with respect to solar distance to obtain power law dependencies. Finally, by combining the found solar cycle and solar distance relations, we obtain a simple dynamical solar wind model for the inner heliosphere, confined to the ecliptic region.}	%methods
{\textcolor{Green4}{ The inclusion of SSN predictions and the extrapolation to the PSP perihelion region enables us to estimate the solar wind environment for PSP's planned trajectory during its prime mission duration 2018--2025.} \hl{The estimated solar wind values at PSP's nearest perihelion are: ... Their value vary up to} \SI{+-0.0000}{\percent}, \hl{arising only from differing amplitude assumptions for the next solar cycle.}}	%results
{}	%conclusions

### 2017-07-16 paperMVVB/sections/content.tex
replaced 'exponential' with 'power law'

%in situ spacecraft
The idea of a space mission flying through the solar corona \hl{dates back to the founding year of NASA in 1958 (ref.).} Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in the case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}}, having even left the heliospause into interstellar space \citep{Gurnett2013}. Ulysses was the unique probe that orbited the Sun out of the ecliptic plane and thus could retrieve solar wind measurements from the poles of the Sun \hl{(ref.)}. So far, Helios~2 made the nearest in situ solar wind measurements ever at a distance as close to the Sun as \SI{0.29}{\au}, closely followed by Helios~1 with \SI{0.31}{\au}.

@@

\color{black}
%frequency distribution data
One cannot know which specific solar wind type or structure PSP will encounter at a given point in time, so we extrapolate the parameters' probability distributions from existing solar wind measurements.

Our approach is to get an analytical representation of the frequency distributions' shapes (Sect.~\ref{sec:frequency_distribution}), their solar activity dependence (Sect.~\ref{sec:solar_activity_variations}) and their solar distance scaling (Sect.~\ref{sec:solar_distance_dependency}). We get the parameters' frequency distributions and solar activity dependence from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles and their distance dependency from solar wind measurements of more than half a solar cycle, covering more than two third of the distance to the Sun (\SIrange{0.29}{0.98}{\au}).%0.68~au\\

@@

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Plot of the velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
For this reason we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

%solar activity = SSN $\neq$ solar cycle\\
The solar wind velocity and its close friends density and temperature are known to depend on the state of the solar cycle \citep{Schwenn1983}, which is why they follow the SSN indirectly (with time lag).
%density modulation with SSN \citep{Schwenn1983} p.~499\\
Thus we derive the correlation coefficients for different time lags between solar wind parameters and SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{Plot of the solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} with cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

@@

%exponential fitting
%%%%%%%%%%%%%%%%%%%%
\section{Solar distance dependency}
\label{sec:solar_distance_dependency}
In this section we use Helios data to obtain exponential fit functions for the heliocentric distance dependency and also evaluate the fits' extrapolation behavior in direction to the Sun. To fit the bulk solar wind distributions' distance dependency we use the frequency fitting method from Sect.~\ref{sec:frequency_distribution} on distance-binned Helios data. This results in models comprising of exponentially with distance shifted lognormal functions.

@@

\subsection{Exponential fitting}
\label{sec:exponential_fitting}
%why exponential function?
An exponential distance behavior is expected from all four parameters (cites?). Therefore we use the exponential function
\begin{align}
	x(r) = d\,r^e	\label{eq:exponential_function}
\end{align}
with the solar distance $r$ for the regression fit of the median and mean. The fits are weighted by data counts per bin.
%fit result table and figure
With $r$ in astronomical units we get the fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) as given in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four parameters from the combined Helios data set. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distance is the point where the fitted median and mean intersect. The year variation is the weighted standard deviation from all yearly fitted exponents.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Year variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.005\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

Those crossings limit the possible extrapolation distances with lognormal functions. To circumvent these limitations for all four solar wind parameters we set the exponents $e_\text{med}$ and $e_\text{avg}$ to be identical, avoiding crossing of median and mean. Then the distribution's width scales exponentially with solar distance. Applying this approximation, we have to accept larger model errors, especially for the magnetic field strength. It also limits the extrapolation accuracy, however it keeps the model simpler.

\subsection{Exponential lognormal fitting}
To retrieve the frequency distributions for every \SI{0.01}{\au} solar distance bin, we choose the same solar wind parameter binning as with the OMNI data (Sect.~\ref{sec:omni_frequency_data}).
% Helios histogram bin sizes for mean of frequency distribution (at specific solar distance):\\
% $B$: bin size 0.5~nT, min 337, mean precision: 0.000545\\
% $v$: bin size 1~cm$^{-3}$, min 497, mean precision: 0.00449\\
% $n$: bin size 10~km/s, min 497, mean precision:  0.00449\\
% $T$: bin size 10\,000~K, min 497, mean precision: 4.49--44.49\\

As mentioned before, we set the exponents of median and average to be identical. Implementing the exponential distance dependency~(\ref{eq:exponential_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get three fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and the common exponent $e'$). Naturally, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and from the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.
\begin{table*}
	\caption{Fit coefficients from the single lognormal exponential function, respectively double lognormal for the velocity (combined Helios data). The errors in brackets are the estimated standard deviations of each fit parameter.}
	\label{tab:extrapolation_model_fit_parameters}
	\centering
	\sisetup{table-figures-integer=1, table-figures-decimal=3, table-figures-exponent=0}
	\begin{tabular}{l c
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = +1.4(2)]
	S[table-format = 1.3(2)]}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Exponent}	&\multicolumn{1}{c}{Balance}\\
			&	&\multicolumn{1}{c}{$d'_\text{med}$}	&\multicolumn{1}{c}{$d'_\text{avg}$}	&\multicolumn{1}{c}{$e'$}	&$c'$\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&5.358(25)	&5.705(28)	&-1.662(11)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Density}	&5.424(33)	&6.845(47)	&-2.114(20)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Temperature}	&6.357(64)	&10.72(14)	&-1.100(20)	&\multicolumn{1}{c}{--}\\
		\hline
		\multirow{1}{*}{Velocity}	&$W''_1$	&3.707(13)	&3.748(16)	&\multirow{2}{*}{0.0990(51)}	&\multirow{2}{*}{0.557(45)}\\
			&$W''_2$	&5.26(13)	&5.42(11)	&	&\\
		\cline{2-6}
		\multicolumn{2}{r}{$W''_\text{II}$}	&4.13(13)\tablefootmark{b}	&4.47(11)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&\multicolumn{1}{c}{--}\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
		\tablefoottext{b}{Velocity median and mean \SI{1}{\au} values for the resulting function. Error estimates derived from the individual fit part errors.}
	}
\end{table*}
% Vdbl fit:
% medians = 370.7, 526
% means = 374.8, 542
% c = 0.557
% via calculate_twolognormal_median.py:
% mean = 446.64
% median = 413.33

@@

The fit models seem to resemble the data quite well (Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}). The magnetic field strength frequency is more focused (around \SI{40}{nT}) at the lower distance boundary than the model's is. This is expected because of our fixed distance independent shape approximation.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Solar wind parameter's frequency distributions over solar distance. Plotted are the binned Helios data and the exponential lognormal fit model (double lognormal for the velocity) with their median values (white).}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The velocity and temperature models' upper values have a higher frequency than the data shows. This is due to the systematic fit discrepancy of the lognormal distribution's high value tails (see zoom box in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}).

@@

%are the exponents time-independent?
Under the assumption that the exponential fall-off laws do not change with time/solar activity (as shown above...), they can be used in general.

We combine the fit coefficients of the median relation for solar activity dependence (\ref{eq:median_with_ssn}) with the ones from the exponential distance dependence (previous section)
\begin{align}
	x_\text{med}(ssn,r) &= (a_\text{med} \cdot ssn + b_\text{med}) \cdot r^{e'}
\end{align}

@@

%magnetic field:
The magnetic field magnitude from our extrapolation is flatter than the analytical magnetic field model from \citet{Banaszkiewicz1998}, who constructed a dipole plus quadrupole plus current sheet (DQCS) model. We attribute this effect to the from a lognormal shape deviating distribution (see Sect.~\ref{sec:exponential_fitting}).\\

>>>>

%in situ spacecraft
The idea of a space mission flying through the solar corona \hl{dates back to the founding year of NASA in 1958 (ref.).} Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in the case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}}, having even left the heliospause into interstellar space \citep{Gurnett2013}. Ulysses was the unique probe that orbited the Sun out of the ecliptic plane and thus could retrieve solar wind measurements from the poles of the Sun \hl{(ref.)}. So far, Helios~2 made the nearest in situ solar wind measurements ever at a distance as close to the Sun as \SI{0.29}{\au}, closely followed by Helios~1 with \SI{0.31}{\au} \citep{Rosenbauer1977}.

@@

\color{black}
Our approach is to get an analytical representation of the frequency distributions' shapes (Sect.~\ref{sec:frequency_distribution}), their solar activity dependence (Sect.~\ref{sec:solar_activity_variations}) and their solar distance scaling (Sect.~\ref{sec:solar_distance_dependency}). We get the parameters' frequency distributions and solar activity dependence from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles and their distance dependency from solar wind measurements of more than half a solar cycle, covering more than two third of the distance to the Sun (\SIrange{0.29}{0.98}{\au}).%0.68~au\\

@@

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{The velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
For this reason we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

%solar activity = SSN $\neq$ solar cycle\\
The solar wind velocity and its close friends density and temperature are known to depend on the state of the solar cycle \citep{Schwenn1983}, which is why they follow the SSN indirectly (with time lag).
%density modulation with SSN \citep{Schwenn1983} p.~499\\
Thus we derive the correlation coefficients for different time lags between solar wind parameters and SSN (see Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{The solar wind parameter yearly medians from OMNI data and the yearly SSN from the \citet{sidc} with cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

@@

%exponential fitting
%%%%%%%%%%%%%%%%%%%%
\section{Solar distance dependency}
\label{sec:solar_distance_dependency}
In this section we use Helios data to obtain power law fit functions for the heliocentric distance dependency and also evaluate the fits' extrapolation behavior in direction to the Sun. To fit the bulk solar wind distributions' distance dependency we use the frequency fitting method from Sect.~\ref{sec:frequency_distribution} on distance-binned Helios data. This results in models comprising of with distance shifted lognormal functions.

@@

\subsection{Power law fitting}
\label{sec:power_law_fitting}
%why power law function?
A power law scaling distance behavior is expected from all four parameters (cites?). Therefore we use the power function
\begin{align}
	x(r) = d\,r^e	\label{eq:power_function}
\end{align}
with the solar distance $r$ for the regression fit of the median and mean. The fits are weighted by data counts per bin.
%fit result table and figure
With $r$ in astronomical units we get the fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) as given in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four parameters from the combined Helios data set. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distance is the point where the fitted median and mean intersect. The year variation is the weighted standard deviation from all yearly fitted exponents.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Year variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.005\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

Those crossings limit the possible extrapolation distances with lognormal functions. To circumvent these limitations for all four solar wind parameters we set the exponents $e_\text{med}$ and $e_\text{avg}$ to be identical, avoiding crossing of median and mean. Then the distribution's width has a power law scaling with solar distance. Applying this approximation, we have to accept larger model errors, especially for the magnetic field strength. It also limits the extrapolation accuracy, however it keeps the model simpler.

\subsection{Power law lognormal fitting}
To retrieve the frequency distributions for every \SI{0.01}{\au} solar distance bin, we choose the same solar wind parameter binning as with the OMNI data (Sect.~\ref{sec:omni_frequency_data}).
% Helios histogram bin sizes for mean of frequency distribution (at specific solar distance):\\
% $B$: bin size 0.5~nT, min 337, mean precision: 0.000545\\
% $v$: bin size 1~cm$^{-3}$, min 497, mean precision: 0.00449\\
% $n$: bin size 10~km/s, min 497, mean precision:  0.00449\\
% $T$: bin size 10\,000~K, min 497, mean precision: 4.49--44.49\\

As mentioned before, we set the exponents of median and average to be identical. Implementing the power law distance dependency~(\ref{eq:power_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get three fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and the common exponent $e'$). Naturally, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and from the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.
\begin{table*}
	\caption{Fit coefficients from the single lognormal power function, respectively double lognormal for the velocity (combined Helios data). The errors in brackets are the estimated standard deviations of each fit parameter.}
	\label{tab:extrapolation_model_fit_parameters}
	\centering
	\sisetup{table-figures-integer=1, table-figures-decimal=3, table-figures-exponent=0}
	\begin{tabular}{l c
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = 1.3(2), table-space-text-post = a, table-align-text-post = false]
	S[table-format = +1.4(2)]
	S[table-format = 1.3(2)]}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Exponent}	&\multicolumn{1}{c}{Balance}\\
			&	&\multicolumn{1}{c}{$d'_\text{med}$}	&\multicolumn{1}{c}{$d'_\text{avg}$}	&\multicolumn{1}{c}{$e'$}	&$c'$\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&5.358(25)	&5.705(28)	&-1.662(11)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Density}	&5.424(33)	&6.845(47)	&-2.114(20)	&\multicolumn{1}{c}{--}\\
		\multicolumn{2}{l}{Temperature}	&6.357(64)	&10.72(14)	&-1.100(20)	&\multicolumn{1}{c}{--}\\
		\hline
		\multirow{1}{*}{Velocity}	&$W''_1$	&3.707(13)	&3.748(16)	&\multirow{2}{*}{0.0990(51)}	&\multirow{2}{*}{0.557(45)}\\
			&$W''_2$	&5.26(13)	&5.42(11)	&	&\\
		\cline{2-6}
		\multicolumn{2}{r}{$W''_\text{II}$}	&4.13(13)\tablefootmark{b}	&4.47(11)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&\multicolumn{1}{c}{--}\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{Values in their respective units \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
		\tablefoottext{b}{Velocity median and mean \SI{1}{\au} values for the resulting function. Error estimates derived from the individual fit part errors.}
	}
\end{table*}
% Vdbl fit:
% medians = 370.7, 526
% means = 374.8, 542
% c = 0.557
% via calculate_twolognormal_median.py:
% mean = 446.64
% median = 413.33

@@

The fit models seem to resemble the data quite well (Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}). The magnetic field strength frequency is more focused (around \SI{40}{nT}) at the lower distance boundary than the model's is. This is expected because of our fixed distance independent shape approximation.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Solar wind parameter's frequency distributions over solar distance. Plotted are the binned Helios data and the power law lognormal fit model (double lognormal for the velocity) with their median values (white).}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The velocity and temperature models' upper values have a higher frequency than the data shows. This is due to the systematic fit discrepancy of the lognormal distribution's high value tails (see zoom box in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}).

@@

%are the exponents time-independent?
Under the assumption that the fall-off laws do not change with time/solar activity (as shown above...), they can be used in general.

We combine the fit coefficients of the median relation for solar activity dependence (\ref{eq:median_with_ssn}) with the ones from the power law distance dependence (previous section)
\begin{align}
	x_\text{med}(ssn,r) &= (a_\text{med} \cdot ssn + b_\text{med}) \cdot r^{e'}
\end{align}

@@

%magnetic field:
The magnetic field magnitude from our extrapolation is flatter than the analytical magnetic field model from \citet{Banaszkiewicz1998}, who constructed a dipole plus quadrupole plus current sheet (DQCS) model. We attribute this effect to the from a lognormal shape deviating distribution (see Sect.~\ref{sec:power_law_fitting}).\\
